---
title: "R analysis test data"
output:
  html_document:
    toc: yes
    theme: united
    print_df: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documentos/Repos/UOC-Master")
```

```{r message=FALSE, warning=FALSE}
library(stringr)
library(reshape2)
library(readr)
library(dplyr)
library(patchwork)
library(ggplot2)
library(knitr)
```

Import CARD data aro index:

```{r message=FALSE}
aro_index <- read_delim("PRJNA307231/aro_index.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

dim(aro_index)
length(unique(aro_index$`ARO Accession`))

y<-colsplit(aro_index$`ARO Accession`, ":", c("X", "ARO Number"))
aro_index<-cbind(aro_index, select(y, `ARO Number`))
```

# Calculate RPKM

## RGI

```{r message=FALSE}
library(readr)
RGI_data <- read_delim("data_testing/RGI_analysis/argannot-150bp-10000-reads.fastq_RGI_output.allele_mapping_data.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
kable(head(RGI_data))
```

Obtain AMR gene family from CARD database:

```{r}
RGI_data<-merge(
select(RGI_data, -`AMR Gene Family`, -`Drug Class`, -`Resistance Mechanism`),
select(aro_index,`Model Sequence ID`, `AMR Gene Family`,`Drug Class`, `Resistance Mechanism`,`Protein Accession`,`ARO Number`, `ARO Name` ), by.x = "ARO Accession", by.y="ARO Number", all.x = TRUE)
```

```{r}
table(is.na(RGI_data$`AMR Gene Family`))
```

Calculate RPKM:

```{r}
library(dplyr)
totalNumReads<-	10000
RGI_data<-RGI_data%>%mutate(RPKM=`All Mapped Reads` /(`Reference Length`/1000 * totalNumReads/1000000))
```

### Average MAPQ distribution

```{r}
library(ggplot2)
ggplot(RGI_data, aes(`Average MAPQ (Completely Mapped Reads)`)) +
  geom_histogram(bins=50, aes(y = stat(density)))+geom_density()+theme_bw()
```

### Apply Average MAPQ filter:

```{r}
RGI_data<-RGI_data%>%filter(`Average MAPQ (Completely Mapped Reads)`>10)
```

### RPKM sum per AMR gene family

```{r}
library(knitr)
n_distinct(RGI_data$`AMR Gene Family`)
AMR_family_RGI<-RGI_data%>%
  group_by(`AMR Gene Family`)%>%
  summarise('number of AMR genes per family (RGI)'=n(),
            sum_RPKM_RGI=sum(RPKM),
            avg_MAPQ=mean(`Average MAPQ (Completely Mapped Reads)`),
            q1_MAPQ=quantile(`Average MAPQ (Completely Mapped Reads)`, 0.25),
            p3_MAPQ=quantile(`Average MAPQ (Completely Mapped Reads)`,0.75),)%>%arrange(desc(sum_RPKM_RGI))
kable(head(AMR_family_RGI))
```

# ARIBA

```{r message=FALSE}
Ariba_data <- read_delim("data_testing/ariba_analysis/argannot-150bp-10000-reads_ariba_results/report.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
```

Remove contig and variant data, select only coding sequences

```{r}
Ariba_data<-Ariba_data %>% filter(gene==1)%>%select(c(1:8,31))%>%unique()

```

```{r eval=FALSE, include=FALSE}
##Coding and no coding sequences
##Ariba_data<-Ariba_data%>%select(c(1:8,31))%>%unique()
```

Calculate RPKM:

```{r}
Ariba_data<-Ariba_data%>%mutate(RPKM=reads /(ref_len/1000 * totalNumReads/1000000))
kable(head(Ariba_data))
```

### Obtain AMR gene family

Trasform ARIBA notation to merge with CARD DB

```{r}
library(stringr)
library(reshape2)
y<-colsplit(Ariba_data$ref_name, "\\.", c("ARO Term", "ARO Accession","NCBI","init_final","Model Sequence ID","X"))
y$NCBI<-ifelse(str_detect(y$init_final, "_"),y$NCBI ,paste(y$NCBI,".",y$init_final) )
y$init_final<-ifelse(str_detect(y$init_final, "_"), y$init_final, y$`Model Sequence ID`)
y$`Model Sequence ID`<-ifelse(str_detect(y$`Model Sequence ID`, "_"), y$X, y$`Model Sequence ID`)
y$`ARO Accession`<-ifelse(y$`ARO Accession`<3000000, y$NCBI, y$`ARO Accession`)
y$`ARO Accession`<-str_split_fixed(y$`ARO Accession`, "\\.", 2)[ ,1]
y<-select(y, -X)
Ariba_data<-cbind(Ariba_data,y)
Ariba_data$`ARO Accession`<-as.numeric(Ariba_data$`ARO Accession`)

Ariba_data<-merge(Ariba_data, select(aro_index,`Model Sequence ID`, `AMR Gene Family`,`Drug Class`, `Resistance Mechanism`,`Protein Accession`,  `ARO Number`, `ARO Name`) , by.x = "ARO Accession", by.y="ARO Number", all.x = TRUE)
```

```{r eval=FALSE, include=FALSE}
table(is.na(Ariba_data$`AMR Gene Family`))
```

### RPKM sum per AMR gene family

```{r}
library(knitr)
library(dplyr)
AMR_family_ARIBA<-Ariba_data%>%
  group_by(`AMR Gene Family`)%>%
  summarise('number of AMR genes per family(ARIBA)'=n(),
            'sum_RPKM_ariba'=sum(RPKM))%>%arrange(desc('sum_RPKM_ariba'))

kable(head(AMR_family_ARIBA))
```

# GROOT

Get groot report (uniqseq)

```{r message=FALSE, warning=FALSE}
library(readr)
Groot_data <- read_delim("data_testing/groot_analysis/argannot-150bp-10000-reads-uniqseq-0report", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)
names(Groot_data)<-c("AMR.gene", "read.count", "gene.length", "coverage.cigar")
kable(head(Groot_data))
```

Trasform GROOT notation and merge with CARD DB

```{r}
library(stringr)
library(reshape2)
library(dplyr)
y<-colsplit(Groot_data$AMR.gene, "\\.", c("ARO Term", "ARO Accession","NCBI","init_final","Model Sequence ID","X"))
y$NCBI<-ifelse(str_detect(y$init_final, "-"),y$NCBI ,paste(y$NCBI,".",y$init_final, sep = "") )
y$init_final<-ifelse(str_detect(y$init_final, "-"), y$init_final, y$`Model Sequence ID`)
y$`Model Sequence ID`<-ifelse(str_detect(y$`Model Sequence ID`, "-"), y$X, y$`Model Sequence ID`)
y$`ARO Accession`<-ifelse(y$`ARO Accession`<3000000, y$NCBI, y$`ARO Accession`)
y$`ARO Accession`<-str_split_fixed(y$`ARO Accession`, "\\.", 2)[ ,1]
y<-select(y, -X)
Groot_data<-cbind(Groot_data,y)
Groot_data$`ARO Accession`<-as.numeric(Groot_data$`ARO Accession`)
Groot_data<-merge(Groot_data, select(aro_index,`Model Sequence ID`, `AMR Gene Family`,`Drug Class`, `Resistance Mechanism`,`Protein Accession`, `ARO Number`, `ARO Name`) ,  by.x = "ARO Accession", by.y="ARO Number", all.x = TRUE)
```

```{r eval=FALSE, include=FALSE}
table(is.na(Groot_data$`AMR Gene Family`))
## 3 secuencias de Groot no mapean con la base de datos CARD: 
## Versión anterior de CARD?
```

Remove non coding AMR sequences:

```{r}
Groot_data<-Groot_data%>%filter(!is.na(`Protein Accession`))
```

Calculate RPKM:

```{r}
Groot_data<-Groot_data%>%mutate(RPKM=read.count /(gene.length/1000 * totalNumReads/1000000))
```

```{r}
library(knitr)
n_distinct(Groot_data$`AMR Gene Family`)
AMR_family_Groot<-Groot_data%>%
  group_by(`AMR Gene Family`)%>%
  summarise('number of AMR genes per family (groot)'=n(),
            sum_RPKM_groot=sum(RPKM))%>%arrange(desc(sum_RPKM_groot))
kable(head(AMR_family_Groot))
```

# ARIBA vs RGI vs GROOT

Number of AMR Gene Families detected by each pipeline:

```{r}
length(AMR_family_ARIBA$`AMR Gene Family`)
length(AMR_family_RGI$`AMR Gene Family`)
length(AMR_family_Groot$`AMR Gene Family`)
```

Merge data obtained from the 3 pipelines:

```{r}
AMR_family<-merge(AMR_family_ARIBA, AMR_family_RGI[ ,1:3], by="AMR Gene Family", all=TRUE)
AMR_family<-merge(AMR_family, AMR_family_Groot, by="AMR Gene Family", all=TRUE)
```

```{r}
venntransform<-function(x){
  x$RGI<-ifelse(is.na(x$sum_RPKM_RGI),FALSE,TRUE)
  x$Ariba<-ifelse(is.na(x$sum_RPKM_ariba),FALSE,TRUE)
  x$Groot<-ifelse(is.na(x$sum_RPKM_groot),FALSE,TRUE)
  x$All<-ifelse(x$RGI==TRUE & x$Ariba==TRUE & x$Groot==TRUE, TRUE, FALSE)
  return(x)}

library(ggvenn)
vennplot<-function(x){
y<-ggplot(x) + geom_venn(aes(A = RGI, B = Ariba, C = Groot),fill_color = c("chartreuse", "orange", "darkorchid1"), text_size=4, set_name_size = 2, stroke_size = 0.2, show_percentage = FALSE)+theme_void()+coord_fixed()+theme(title=element_text(size=5))
  return(y)}
```

```{r}
AMR_family_venn<-venntransform(AMR_family)
vennplot(AMR_family_venn)
```

Transform data according to which pipeline has detected each AMR gene family:

```{r}
data_trans<-function(x){
x$RGI<-ifelse(is.na(x$sum_RPKM_RGI),0,1)
x$Ariba<-ifelse(is.na(x$sum_RPKM_ariba),0,1)
x$Groot<-ifelse(is.na(x$sum_RPKM_groot),0,1)
x$npipelines<-apply(select(x, RGI, Ariba,Groot),1, sum)
x$npipelines<-as.factor(x$npipelines)
x$pipelines<-ifelse(x$npipelines==3, "All", 
                                ifelse(x$npipelines==2 & x$RGI==0, "Groot+Ariba",
                                       ifelse(x$npipelines==2 & x$Ariba==0, "RGI+Groot",
                                              ifelse(x$npipelines==2 & x$Groot==0, "RGI+Ariba",
                                                     ifelse(x$RGI==1, "RGI only",
                                                            ifelse(x$Ariba==1, "Ariba only", "Groot only"))))))
x$pipelines_class<-ifelse(x$npipelines==1,0,1)
return(x)}
```

```{r}
AMR_family<-data_trans(AMR_family)
```

Store discordant families names:

```{r}
discrodant_families<-list(AMR_family[AMR_family$pipelines=="RGI only" ,1],
AMR_family[AMR_family$pipelines=="Groot only" ,1],
AMR_family[AMR_family$pipelines=="RGI+Groot" ,1],
AMR_family[AMR_family$pipelines=="Groot+Ariba" ,1])
names(discrodant_families)<-c("RGI only", "Groot only", "RGI+Groot", "Groot+Ariba")
```

```{r}
capture.output(discrodant_families, file = "discrodant_families.txt")
```

```{r eval=FALSE, include=FALSE}
library(cutpointr)

cp_RGI<-AMR_family %>% filter(RGI==1) %>% cutpointr(sum_RPKM_RGI, pipelines_class)
##cp_ariba<-AMR_family %>% filter(Ariba==1) %>% cutpointr(sum_RPKM_ariba, pipelines_class) 
##Class ariba only does not exist
##cp_groot<-AMR_family %>% filter(Groot==1) %>% cutpointr(sum_RPKM_groot, pipelines_class)

ROC<-function(cp){
plot_roc(cp)+theme_classic()+
geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dotted")+
annotate("text", x = 0.5, y = 0.25, label=paste("Value:",round(cp$optimal_cutpoint,2), "\n Sens:", round(cp$sensitivity,2), "\n Spe:", round(cp$specificity,2)), hjust=0)
  }

roc_RGI<-ROC(cp_RGI)+ggtitle("ROC curve RGI")
```

RPKM distribution of AMR gene families according to the number of pipelines:

```{r}
AMR_family$pipelines<-factor(AMR_family$pipelines, ordered = TRUE, levels = c("RGI only", "Ariba only", "Groot only", "Groot+Ariba", "RGI+Groot", "RGI+Ariba", "All"))

p_RGI<-ggplot(subset(AMR_family, AMR_family$RGI==1), aes(pipelines, sum_RPKM_RGI))+geom_boxplot(aes(colour=pipelines), outlier.shape = NA, na.rm = TRUE)+geom_jitter(width = 0.2, aes(colour=pipelines))+theme_classic()+theme(legend.position = "none")+ggtitle("RGI")+scale_y_continuous(trans='log10')+theme(axis.text.x = element_text(angle=45, hjust = 1), axis.title.x = element_blank())

p_Ariba<-ggplot(subset(AMR_family, AMR_family$Ariba==1), aes(pipelines, sum_RPKM_ariba))+geom_boxplot(aes(colour=pipelines), outlier.shape = NA, na.rm = TRUE)+geom_jitter(width = 0.2, aes(colour=pipelines))+theme_classic()+theme(legend.position = "none")+ggtitle("Ariba")+scale_y_continuous(trans='log10')+theme(axis.text.x = element_text(angle=45, hjust=1), axis.title.x = element_blank())

p_Groot<-ggplot(subset(AMR_family, AMR_family$Groot==1), aes(pipelines, sum_RPKM_groot))+geom_boxplot(aes(colour=pipelines), outlier.shape = NA, na.rm = TRUE)+geom_jitter(width = 0.2, aes(colour=pipelines))+theme_classic()+theme(legend.position = "none")+ggtitle("Groot")+scale_y_continuous(trans='log10')+theme(axis.text.x = element_text(angle=45, hjust=1),axis.title.x = element_blank())
p_RGI+p_Ariba+p_Groot
```

```{r eval=FALSE, include=FALSE}
##write_tsv(AMR_family, "AMR_family.tsv")
```

### Discordance analysis:

The pipelines variable (number of pipelines that have detected an AMR gene family) is added to each pipeline report:

```{r}
RGI_data<-merge(RGI_data, select(AMR_family, pipelines, `AMR Gene Family`), by="AMR Gene Family", all.x = TRUE)

Groot_data<-merge(Groot_data, select(AMR_family, pipelines, `AMR Gene Family`), by="AMR Gene Family", all.x = TRUE)

Ariba_data<-merge(Ariba_data, select(AMR_family, pipelines, `AMR Gene Family`), by="AMR Gene Family", all.x = TRUE)
```

```{r eval=FALSE, include=FALSE}
table(RGI_data$pipelines)
```

For each pipeline the discordant sequences are selected:

```{r}
RGI_seqs<-RGI_data%>%filter(pipelines!="All")%>%select(`Reference Sequence`)
Groot_seqs<-Groot_data%>%filter(pipelines!="All")%>%select(AMR.gene)
Ariba_seqs<-Ariba_data%>%filter(pipelines!="All")%>%select(ref_name)
```

```{r eval=FALSE, include=FALSE}
write.table(RGI_seqs, "data_testing/RGI_analysis/RGI_seqs.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(Groot_seqs, "data_testing/groot_analysis/Groot_seqs.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
write.table(Ariba_seqs, "data/testing/ariba_analysis/Ariba_seqs.txt", quote = FALSE, row.names = FALSE, col.names = FALSE)
```

The original sequences belonging to discordant AMR gene families for each pipeline are selected (BASH):

```{bash}
cd data_testing/RGI_analysis
samtools view argannot-150bp-10000-reads.fastq_RGI_output.temp.bam | grep -f RGI_seqs.txt > RGI_seqs_bam.txt
```

```{bash}
cd data_testing/groot_analysis
samtools view argannot-150bp-10000-reads-uniqseq.bam | grep -f Groot_seqs.txt > Groot_seqs_bam.txt
```

Ariba does not generate any BAM file.

Open the generated files:\
data_seq=sample sequence name\
Reference Sequence= Reference sequence name

```{r message=FALSE, warning=FALSE}
library(readr)
RGI_seqs_bam <- read_delim("data_testing/RGI_analysis/RGI_seqs_bam.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)
RGI_seqs_bam<-RGI_seqs_bam%>%select(X1, X3)%>%rename(data_seq=X1, `Reference Sequence`=X3 )
head(RGI_seqs_bam)
```

```{r message=FALSE, warning=FALSE}
Groot_seqs_bam <- read_delim("data_testing/groot_analysis/Groot_seqs_bam.txt", 
    "\t", escape_double = FALSE, col_names = FALSE, 
    trim_ws = TRUE)
Groot_seqs_bam<-Groot_seqs_bam%>%select(X1, X3)%>%rename(data_seq=X1, `Reference Sequence`=X3 )
Groot_seqs_bam$`Reference Sequence`<-str_remove(Groot_seqs_bam$`Reference Sequence`, "\\*")
head(Groot_seqs_bam)
```

Merge data to obtain CARD information from Reference sequences and extract a simpler version of the sample sequence names. Compare sample sequence names to ARO Term names from Reference sequences.

```{r}
RGI_seqs_bam<-merge(RGI_seqs_bam, select(RGI_data, `AMR Gene Family`,`Reference Sequence`,`ARO Term`, `Resistance Mechanism`, `Drug Class`, RPKM, pipelines), by="Reference Sequence", all.x = TRUE)
RGI_seqs_bam$data_seq_simplified<-str_split_fixed(RGI_seqs_bam$data_seq, "~~~", 3)[,2]
RGI_seqs_bam$data_seq_simplified<-str_split_fixed(RGI_seqs_bam$data_seq_simplified, "\\)", 2)[ ,2]
RGI_seqs_bam$concordance<-ifelse(tolower(RGI_seqs_bam$data_seq_simplified)%in%tolower(RGI_seqs_bam$`ARO Term`) | tolower(RGI_seqs_bam$`ARO Term`)%in%tolower(RGI_seqs_bam$data_seq_simplified),1,0)
```

```{r}

Groot_seqs_bam<-merge(Groot_seqs_bam, select(Groot_data, `AMR Gene Family`,AMR.gene,`ARO Term`, `Resistance Mechanism`, `Drug Class`, RPKM, pipelines), by.x="Reference Sequence", by.y="AMR.gene", all.x = TRUE)
Groot_seqs_bam$data_seq_simplified<-str_split_fixed(Groot_seqs_bam$data_seq, "~~~", 3)[,2]
Groot_seqs_bam$data_seq_simplified<-str_split_fixed(Groot_seqs_bam$data_seq_simplified, "\\)", 2)[ ,2]
Groot_seqs_bam$concordance<-ifelse(tolower(Groot_seqs_bam$data_seq_simplified)%in%tolower(Groot_seqs_bam$`ARO Term`) | tolower(Groot_seqs_bam$`ARO Term`)%in%tolower(Groot_seqs_bam$data_seq_simplified) ,1,0)
```

Manual revision of concordance:

```{r}
write_tsv(RGI_seqs_bam, "data_testing/RGI_analysis/RGI_seqs_bam.tsv")
write_tsv(Groot_seqs_bam, "data_testing/groot_analysis/Groot_seqs_bam.tsv")
```

## Sensitivity and Specificity

True positives: concordant sequences (identified by 3 pipelines) + correctly identified discordant sequences False positives: incorrectly identified discordant sequences False negatives: not mapped sequences True negatives: 0, all sequences are resistant genes.

Total: 10000 seq

```{r message=FALSE, warning=FALSE}
RGI_seqs_bam<- read_delim("data_testing/RGI_analysis/RGI_seqs_bam_manual.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
nrow(RGI_seqs_bam)
table(RGI_seqs_bam$concordance)
```

### RGI

-   9473 mapped sequences: (mapping stats)

    -   232 discordant sequences

        -   224 correctly identified\
        -   8 incorrectly identified\

    -   9241 concordant sequences (9473-232)\

-   527 not mapped sequences

```{r}
vp<-9241+232
fp<-8
fn<-527
vn<-0
sensibilidad_rgi<-vp*100/(vp+fn)
especificidad_rgi<-vn*100/(vn+fp)
vpp_rgi<-vp*100/(vp+fp)

sensibilidad_rgi
especificidad_rgi
vpp_rgi
```

### GROOT

```{bash}
samtools view data_testing/groot_analysis/argannot-150bp-10000-reads-uniqseq.bam | wc -l
```

```{r message=FALSE, warning=FALSE}
Groot_seqs_bam<- read_delim("data_testing/groot_analysis/Groot_seqs_bam_manual.csv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
nrow(Groot_seqs_bam)
table(Groot_seqs_bam$concordance)
```

-   8822 mapped sequences: (bam file)

    -   694 discordant sequences

        -   680 correctly identified\
        -   14 incorrectly identified\

    -   8128 concordant sequences (8822-694)

-   1178 not mapped sequences

```{r}
vp<-8128+680
fp<-14
fn<-1178
vn<-0
sensibilidad_groot<-vp*100/(vp+fn)
especificidad_groot<-vn*100/(vn+fp)
vpp_groot<-vp*100/(vp+fp)

sensibilidad_groot
especificidad_groot
vpp_groot
```
